---
title: Twitter 
author: Joel Soroos
date: '2020-12-06'
slug: twitter
categories: []
tags:
  - tidyverse
  - r
  - rstats
  - gt
  - rtweet
  - ggmap
---


## Purpose
The goal of this post is to source and analyze Twitter posts and followers for a given Twitter account ("handle") using R.  We will identify tweets with the most likes and retweets, as well as posting trends over time.  We will also calculate the geographic distribution of Twitter followers by enriching the raw Twitter data with Google Maps API attributes.  

## 1. Setup
First, we need to obtain and authorize a free Twitter Developer account (instructions [here](https://cran.r-project.org/web/packages/rtweet/vignettes/auth.html)).

Second step is to install and load the R [rtweet](https://cran.r-project.org/web/packages/rtweet/rtweet.pdf) package by Michael Kearney. 

We will extract Twitter data multiple times so I have created two variables to avoid duplication and inconsistencies.  The "twitter_user" variable stores the Twitter handle.  The "sample" variable specifies the number of extracted records as some Twitter accounts have large volumes of followers and tweets.

```{r , warning = TRUE, results = FALSE, message = FALSE}
   
   library (rtweet)

   twitter_user <- "@blackyellow"
```


## 2. Analyze tweets
### Source tweets
Now we will extract our first set of Twitter data, which will be tweets from the Borussia Dortmund Football ("BVB") club English account.

The get_timeline function contains 90 fields for each tweet.  

The BVB account mosts frequently so we will limit to the last 200 tweets.

```{r , warning = FALSE, message = FALSE}

   tweets_raw <- get_timeline (twitter_user, n = 500)
```


### Transform Tweets
Many fields are not needed for this exercise so we can cull the dataset.  I have also renamed several fields for brevity and more standard Twitter notation (i.e. likes instead of favorites).

We are focusing on original tweets so have filtered out retweets.

I have extracted year from the creation date to enable summarizing tweet volume by year.  Also, I have truncated the Tweet text field from 280 to 60 characters so more cleanly fits in a formatted table. 

```{r , warning = TRUE, message = FALSE}

   library (tidyverse)
   library (lubridate)

   tweets <- tweets_raw %>%
      filter (is_retweet == FALSE) %>%
      rename (
         likes = favorite_count,
         retweets = retweet_count,
         created = created_at
         ) %>%
      mutate (
         text = str_sub(text,1,60),
         created = as.Date(created),
         year = year(created)
         ) %>%
      rename(date = created) %>%
      select (text, date, year, likes, retweets)
```


### Visualize most liked tweets 
We will sort by most liked tweets over time and rank via the dplyr row_to_column formula.

The [gt package](https://gt.rstudio.com/) by Richard Iannone formats tables in a easy-to-read manner.  The core gt::gt function adds lines between rows.  Many other table formatting enhancements are available but out of scope for this blog.

The tweet with the most likes was BVB's tribute to football legend Diego Maradona, who recently passed away.

```{r chart_likes, warning = TRUE, message = FALSE}

   library (gt)

   tweets %>%
      arrange (-likes, -retweets) %>%
      head (10) %>%
      rowid_to_column("rank") %>%
      select (-year) %>%
      relocate (rank) %>%
      gt () 
```

### Visualize most retweeted tweets
Next we will re-sort and re-rank by retweets, again using the GT package to cleanly format.

A tweet celebrating player Giovanni Reyna signing a contract extension earned the most retweets.  The tribute to Diego Mardona (most liked tweet) was #2 when ranking by retweets.
```{r chart_retweet, warning = TRUE, message = FALSE}

   tweets %>%
      arrange (-retweets, -likes) %>%
      head (10) %>%
      rowid_to_column("rank") %>%
      select (rank, text, date, retweets, likes) %>%
      gt ()
```


### Visualize number of tweets trends
Another area of interest is how tweet activity has evolved over time.  Using dplyr we can easily summarize by number of tweets by day.

The volume of tweets was smallest during the international break when BVB does not play games.  Its players are representing their separate national teams.
```{r warning = F, message =F}

   tweets %>%
      group_by (date) %>%
      count () %>%
      ggplot (aes (x = date, y = n)) +
         geom_col () +
         geom_label (aes(label = n)) +
         theme_minimal() +
         labs (x="", y = "Number of tweets")
```


## 3. Analyze followers
Learning about followers for a Twitter handle is also frequently useful.

Two rtweet package functions are helpful here.  First, get_followers returns a vector of Twitter user_ids of followers.  The lookup_users function then extracts user attributes such as screen name, physical location (if supplied in profile) and number of  followers.  

We will also add a row ID ranking each follower by largest number of followers via dplyr's rowid_to_column function.

```{r warning = F, message =F}

   library(skimr)

  followers <- twitter_user %>%
      get_followers() %>%
      head (50) %>%
      pull (user_id) %>%
      lookup_users () %>%
      select (screen_name, name, location, followers_count) %>%
      arrange (-followers_count) %>%
      rowid_to_column("rank")

   skim (followers)
```


### Top followers
We will first list the 10 largest followers by respective number of followers.
```{r warning = F, message =F}

   followers %>%
      head (10) %>%
      select (rank, name, location, followers_count) %>%
      gt () %>%
         fmt_number (columns = vars(followers_count), use_seps = T, decimals = 0)
```


### Obtain follower geographic attributes
Another interesting item to research on followers is geographic distribution.

The provided location in Twitter profile is a good starting point.  However, locations are neither standardized by level of detail (city, country) nor spelling.  Standardization is necessary to group in table or map.

We will standardize at the highest level of detail, which is country, via the [Google Maps API](https://developers.google.com/maps/documentation) and accessed via the [ggmaps package](https://cran.r-project.org/web/packages/ggmap/ggmap.pdf).  If this is the first time you have accessed the Google Maps API, you will need to complete a one-time step of obtaining a free Google maps API key (instructions [here](https://rdrr.io/cran/ggmap/man/register_google.html)).

```{r warning = F, message =F}
   
   library (ggmap)

   register_google(key = Sys.getenv("GOOGLE_MAPS_API"))

   followers_geo <- followers %>%
      head (sample) %>%
      mutate_geocode (location, output = "more") %>%
      mutate (
         country = word (address, -1,-1),
         country = str_to_upper(country),
         country = ifelse(is.na(country), "NOT LISTED", country)
         ) %>%
      select (rank:followers_count, country, lon, lat) 
   
   followers_geo
```


### Visualize followers by country
We will first identify the top follower countries and summarize in a GT table.
```{r warning = F, message =F}

   followers_agg <- followers_geo %>%
      count (country) %>%
      arrange (-n)
   
   ggplot (followers_agg, aes (x = reorder(country, n), y = n)) +
      geom_col () +
      geom_label (aes(label = n)) +
      coord_flip () +
      theme_minimal() +
      labs (x="", y = "Followers")
```


### Visualize followers on a map
Next we will convert to a map to enable viewing the location of all users along with countries with most users.
```{r warning = F, message =F}
   
   library (maps)

   world <- map_data ("world")

   ggplot () +
      geom_map (data = world, map = world, aes (long, lat, map_id = region), fill = "lightgray", color = "black", size =0.05) +
      geom_point (data = followers_geo, aes (x = lon, y = lat), color = "blue", size = 0.75) + 
      theme_grey ()
```


## 4. Potential future analyses
We are only scratching the surface of available Twitter data, subsequent data enrichment from complementary packages and subsequent analytics.  Second generation followers (followers of followers) could easily be sourced and subtotaled to identify possile impact from a tweet.   Or Twitter users that the account follows could also be extracted.  


## 5. Conclusion 
R offers many powerful tools to source, analyze and visualize Twitter information.  The Twitter API (accessed via the rtweet package) enables free sourcing of follower and tweet attributes that can be easily imported into R.  The raw data can then be enriched by sourcing geographic attributes from Google's API via the ggpmap package.   Tidyverse package such as dplyr, ggplot and GT can then be leveraged to transform and visualize for insight.